### 2020-06-17 report/sections/server/react.tex
Report: feedback

\chapter{\fancyname{ReactMPST}: Front-End Session Type Web Development}
\label{chap:react}

In this chapter, we present \fancyname{ReactMPST},
our session type API generation strategy for browser-side endpoints
implemented using React \cite{React}.
We will refer to the FSM of the \trole{Client}
endpoint (\cref{fig:adderclientfsm})
from the \tprotocol{Adder} protocol 
throughout this chapter.

\begin{figure}[!b]
\centering
\includegraphics[width=0.55\textwidth]{AdderClientFSM}
\captionof{figure}{\trole{Client} Endpoint FSM
in \tprotocol{Adder} protocol}
\label{fig:adderclientfsm}
\end{figure}

@@

\section{Summary}
\begin{itemize}
\item static guarantees on avoiding communication mismatch
\item static guarantees on affine channel usage
\item leverage the react framework to express the model types approach from Fowler 2020 appropriately, to provide static guarantees, by construction, that the channel actions made available at each state respect the protocol, and by definition, multiparty session type safety
\end{itemize}

>>>>

\chapter{\reactcodegen: Front-End Session Type Web Development}
\label{chap:react}

In this chapter, we present \reactcodegen,
our session type API generation strategy for browser-side endpoints
implemented using React \cite{React}.
We will refer to the FSM of the \trole{Client}
endpoint (\cref{fig:adderclientfsm})
from the \tprotocol{Adder} protocol 
throughout this chapter.

We highlight the challenges of session-typed GUI programming
(\cref{section:reactchallenges}) and motivate 
our approach (\cref{section:reactapproach})
for adapting the proposal from
\cite{MVU2020} for the React.js framework, along with a brief
introduction to the framework.
We proceed to explain how we leverage React
to encode the EFSM (\cref{section:reactefsm})
and implement the session runtime (\cref{section:reactruntime}),
focusing on our novel approach to preventing
channel reuse in GUI programming using React.
We conclude by analysing some alternative designs
(\cref{section:reactalt})
and evaluating the limitations of our approach
(\cref{section:reactlimitations}).

\begin{figure}[!b]
\centering
\includegraphics[width=0.6\textwidth]{AdderClientFSM}
\captionof{figure}{\trole{Client} Endpoint FSM
in \tprotocol{Adder} protocol}
\label{fig:adderclientfsm}
\end{figure}

@@

\section{Summary}
In this chapter, we have presented
our session type API generation strategy
which targets the React framework for browser-side endpoints.

We motivated our approach of
implementing the concept of model types introduced in \cite{MVU2020}
using a popular front-end framework in React.
We define an interpretation of model types
using the EFSM obtained from a Scribble protocol.

We discussed how to prevent channel reuse in
GUI programming in React. We build upon the design choices
implemented in \nodecodegen by concealing channel resources
in the session runtime and exposing seams for the developer
to customise what interface to render at each EFSM state,
what user interface events are allowed to trigger a send action,
and how to handle messages received from the channel.

% Developers that implement...

### 2020-06-17 report/sections/server/react/approach.tex
Report: feedback

\section{Approach}
We motivate our approach from \cite{MVU2020} by extending
their work on \textit{multiple model types} motivated by the
\textit{Model-View-Update} architecture (MVU),
introduced in \cref{section:bgrelated}.
The concept of model types express type dependencies between these
components: a \emph{model type} uniquely defines a \textit{view function},
set of \textit{messages} and \textit{update function} -- rather than
producing a new model, the update function defines valid transitions to
other model types.

@@

We implement model types for the EFSM on top of the 
\emph{React.js} framework developed by Facebook \cite{React}.
React is widely used in industry to create scalable single-page
TypeScript applications, so this makes our workflow beneficial in an
industrial context. 
The framework defines a way for data to flow
between UI elements, and empowers the UI to subscribe and
``react'' to data changes.
We aim to implement similar behaviour with respect to the EFSM:
the UI should react to EFSM state transitions,
so we can \textbf{statically} ensure that the
channel actions ``present'' on the browser at any given time
are those permitted by the current EFSM state.
We introduce the framework in \cref{subsection:react}.

@@

\item 
\textbf{S[40-43].tsx:}
Developer APIs for implementing EFSM states 
(\cref{section:reactefsm});

>>>>

\section{Approach}
\label{section:reactapproach}

We motivate our approach from \cite{MVU2020} by extending
their work on \textit{multiple model types} motivated by the
\textit{Model-View-Update} architecture (MVU),
introduced in \cref{section:bgrelated}.
The concept of model types express type dependencies between these
components: a \emph{model type} uniquely defines a \textit{view function},
set of \textit{messages} and \textit{update function} -- rather than
producing a new model, the update function defines valid transitions to
other model types.

@@

We implement model types for the EFSM on top of the 
\emph{React.js} (React) framework developed by Facebook \cite{React}.
React is widely used in industry to create scalable single-page
web applications, so this makes our workflow beneficial in an
industrial context. 
The framework defines a way for data to flow
between UI elements, and empowers the UI to subscribe and
``react'' to data changes.
We aim to implement similar behaviour with respect to the EFSM:
the UI should react to EFSM state transitions,
so we can \textbf{statically} ensure that the
channel actions ``present'' on the browser at any given time
are those permitted by the current EFSM state.
We introduce the framework in \cref{subsection:react}.

@@

\item 
\textbf{S[40-43].tsx\footnote{
The \filename{.tsx} file extension allows
for embedding \textit{JSX} \cite{JSX} elements inside the file.
JSX is a XML-like syntax extension to JavaScript
for elements and components.
}:}
Developer APIs for implementing EFSM states 
(\cref{section:reactefsm});

### 2020-06-17 report/sections/server/react/challenges.tex
Report: progress on chapter 5 limitations
Report: feedback

\section{Challenges}
Our goal with \fancyname{ReactMPST} is to
generate session type implementations for the web browser.
Session typing the user interface (UI) logic is inherently difficult --
assuming channel actions are bound to user interface events 
(i.e. clicking a button sends a message), how does one formalise
channel linearity? How do we guarantee that, if the button
triggers a send at some EFSM state, that it triggers not more than
one channel action, \textit{and} the user cannot trigger the action
at another EFSM state?

Existing work (\cref{subsection:sessiontypewebdev})
do tackle browser-side session typing.
Fowler introduced the concept of model types to prevent
channel linearity violation in his proposal for
integrating session types with
graphical user interface (GUI) programming \cite{MVU2020}, but
it requires the linear type system available from
the proprietary Links web programming language \cite{LINKS}.
The session type-safe web development framework presented
by King et al. \cite{PureScript2019} generates APIs for
a functional target language in PureScript, and relies on
the Concur UI framework that constructs UIs sequentially.

We find these proposals to come at the cost of
limiting developer productivity by adopting unconventional practices
that may require a learning curve.
Through our work, we aim to distil the key
concepts from existing work that provide session type safety
for web-based GUI programming, and implement them using
mainstream front-end web development tools (e.g. TypeScript
and React), to provide developers with an intuitive way to implement
browser-side endpoints that guarantee communication safety.

>>>>

\section{Challenges}
\label{section:reactchallenges}

Our goal with \fancyname{ReactMPST} is to
generate session type implementations for the web browser.
Integrating session types into user interfaces is inherently difficult --
assuming channel actions are bound to user interface events 
(i.e. clicking a button sends a message), how does one formalise
channel linearity? How do we guarantee that, if the button
triggers a send at some EFSM state, that it triggers not more than
one channel action, \textit{and} the user cannot trigger the action
at another EFSM state?

We recap how existing work 
(as discussed in \cref{subsection:sessiontypewebdev})
tackle browser-side session typing.
Fowler introduced the concept of model types to prevent
channel linearity violation in his proposal for
integrating session types with
graphical user interface (GUI) programming \cite{MVU2020}, but
it uses the Links web programming language \cite{LINKS} which
lacks compatibility with the ecosystem of JavaScript
libraries that one might use in front-end development as well.
The session type-safe web development framework presented
by King et al. \cite{PureScript2019} generates APIs for
a functional target language in PureScript, and relies on
the \textit{Concur UI} framework that constructs UIs sequentially.

As motivated in \cref{section:intro},
we find these proposals to come at the cost of
limiting developer productivity by adopting unconventional practices
that may require a learning curve.
Through our work, we aim to distil the key
concepts from existing work that provide session type safety
for web-based GUI programming, and implement them using
mainstream front-end web development tools (i.e. TypeScript
and React), to provide developers with an intuitive way to implement
browser-side endpoints that guarantee communication safety.

### 2020-06-17 report/sections/server/react/efsm.tex
Report: feedback

We encode each EFSM state as a React component.
This encoding is consistent with the semantics of model types,
as outlined in \cref{table:efsmmodeltypes}.
\fancyname{ReactMPST} generates an abstract React component class
for each EFSM state. The developer implements the generated API
by extending the abstract class to define their own
view function and implement any abstract methods
required by the EFSM state.

@@

Without \fancyname{ReactMPST},
an implementation of state S40 from \cref{fig:adderclientfsm}
may resemble the following, where the developer attaches
event listeners on UI elements to trigger channel actions:

@@

\begin{figure}[!h]
\begin{lstlisting}[language=javascript,tabsize=2]
type State = { num1: number, num2: number, message: string }; (*@\label{line:clientsendstatestate}@*)
class InputWindow extends S40<State> {
	render() {
		const { num1, num2, message } = this.state;
		const (*@\hl{Add}@*) = this.ADD('onClick', (*@\label{line:clientsendfactory1}@*)
			(ev: React.MouseEvent) => [num1, num2]); (*@\label{line:clientsendpayload1}@*)
		const (*@\hl{Quit}@*) = this.QUIT('onMouseOver', (*@\label{line:clientsendfactory2}@*)
			(ev: React.MouseEvent) => [message]); (*@\label{line:clientsendpayload2}@*)
		return (<div>
			// Omitting <input>s for entering numbers, message
			(*@\hl{<Add>}@*)<button>Sum</button>(*@\hl{</Add>}@*)
			(*@\hl{<Quit>}@*)<button>Quit</button>(*@\hl{</Quit>}@*)
		</div>);
	}
}
\end{lstlisting}
\captionof{lstlisting}{Developer Implementation for \trole{Client}
Send State in \tprotocol{Adder} protocol}
\label{lst:adderclientsenddev}
\end{figure}

@@

\begin{figure}[!h]
\begin{lstlisting}[language=javascript,tabsize=2,title=S40.tsx]
type Props = { factory: SendComponentFactoryFactory };
abstract class S40<State> extends React.Component<Props, State> {
	protected ADD: SendComponentFactory<[number, number]>;
	protected QUIT: SendComponentFactory<[string]>;

	constructor(props: Props) {
		super(props);
		this.ADD = props.factory<[number, number]>( (*@\label{line:factory1}@*)
			'ADD', ReceiveState.S42);
		this.QUIT = props.factory<[string]>( (*@\label{line:factory2}@*)
			'QUIT', ReceiveState.S43);
	}
}
\end{lstlisting}
\captionof{lstlisting}{Generated Type for \trole{Client} Send State
in \tprotocol{Adder} protocol}
\label{lst:reactsendcomponent}
\end{figure}

@@

According to the EFSM, a receive state means
the endpoint is ``waiting for the message'',
and the actual receive event represents a transition
to a successor state.
This means that the developer's implementation will define
what to render on the DOM when the endpoint is waiting
for the message, rather than updating the DOM with the
received message.
Referring to \cref{fig:adderclientfsm} again,
S42 will be rendered when the endpoint is waiting for
\tmsg{RES}, but once received, S40 will be immediately
rendered. The developer will need to manage the flow
of stateful application data between UI components separately
-- \fancyname{ReactMPST} only generates code for managing
the communication automaton.

@@

Receive states in \fancyname{ReactMPST}
work in a similar way as with \fancyname{NodeMPST}.
The runtime will invoke the \tmsg{RES} method
implemented by the developer upon receiving the message
through the WebSocket.
In order to do so, the S42 component 
needs to \textit{register}
the handler with the runtime.

@@

This raises questions about the type signature for the
\texttt{register} function, as the receive handler
identifiers and type signatures differ between receive states.
We adopt the same strategy as with \fancyname{NodeMPST}:
the receive handler registered by the receive state
to the runtime also performs the message parsing,
so all receive states register a ``general handle function'' typed

@@

\begin{figure}[!h]
\begin{lstlisting}[language=javascript,tabsize=2]
type Props = {
	register: (handle: ReceiveHandler) => void
};
abstract class S42<S> extends React.Component<Props, S> {
	abstract THANKS(payload1: string): void;
	abstract TERMINATE(): void;

	componentDidMount() {
		this.props.register(this.handle.bind(this));
	}

@@

\subsection{Terminal States}
To be consistent with the conventions defined for \fancyname{NodeMPST},
the browser endpoint proactively closes the connection at its terminal state.
We generate an abstract React component for terminal states 
to allow the developer to customise what to
\texttt{render} upon protocol termination.

>>>>

We encode each EFSM state as a React component.
This encoding is consistent with the semantics of model types,
as outlined in \cref{table:efsmmodeltypes}.
\reactcodegen generates an abstract React component class
for each EFSM state. The developer implements the generated API
by extending the abstract class to define their own
view function and implement any abstract methods
required by the EFSM state.

@@

Without \reactcodegen,
an implementation of state S40 from \cref{fig:adderclientfsm}
may resemble the following, where the developer attaches
event listeners on UI elements to trigger channel actions:

@@

\begin{figure}[!h]
\begin{lstlisting}[language=javascript,tabsize=2]
// UI component keeps track of data entered by user
type State = { num1: number, num2: number, message: string }; (*@\label{line:clientsendstatestate}@*)

class InputWindow extends S40<State> {
	render() {
		const { num1, num2, message } = this.state;

		// Generate React component that, when clicked,
		// sends ADD message with `num1` and `num2`.
		const (*@\hl{Add}@*) = this.ADD('onClick', (*@\label{line:clientsendfactory1}@*)
			(ev: React.MouseEvent) => [num1, num2]); (*@\label{line:clientsendpayload1}@*)

		// Generate React component that, when hovered,
		// sends QUIT message with `message` string.
		const (*@\hl{Quit}@*) = this.QUIT('onMouseOver', (*@\label{line:clientsendfactory2}@*)
			(ev: React.MouseEvent) => [message]); (*@\label{line:clientsendpayload2}@*)
		return (<div>
			// Omitting <input>s for entering numbers, message
			(*@\hl{<Add>}@*)<button>Sum</button>(*@\hl{</Add>}@*)
			(*@\hl{<Quit>}@*)<button>Quit</button>(*@\hl{</Quit>}@*)
		</div>);
	}
}
\end{lstlisting}
\captionof{lstlisting}{Developer Implementation for \trole{Client}
Send State in \tprotocol{Adder} protocol}
\label{lst:adderclientsenddev}
\end{figure}

@@

\begin{figure}[!h]
\begin{lstlisting}[language=javascript,tabsize=2,title=S40.tsx]
// Expect to receive `factory` function passed in by runtime
type Props = { factory: SendComponentFactoryFactory };

abstract class S40<State> extends React.Component<Props, State> {

	// Allow developer implementation subclass to use factories
	protected ADD: SendComponentFactory<[number, number]>;
	protected QUIT: SendComponentFactory<[string]>;

	constructor(props: Props) {
		super(props);

		// Generate a factory that generates React components
		// which sends an [ADD] message along with two numbers,
		// then transitions to State S42.
		this.ADD = props.factory<[number, number]>( (*@\label{line:factory1}@*)
			'ADD', ReceiveState.S42);

		this.QUIT = props.factory<[string]>( (*@\label{line:factory2}@*)
			'QUIT', ReceiveState.S43);
	}
}
\end{lstlisting}
\captionof{lstlisting}{Generated Type for \trole{Client} Send State
in \tprotocol{Adder} protocol}
\label{lst:reactsendcomponent}
\end{figure}

@@

According to the EFSM, a receive state means
the endpoint is ``waiting for the message'',
and the actual receive event represents a transition
to a successor state.
This means that the developer's implementation will define
what to render on the DOM when the endpoint is waiting
for the message, rather than updating the DOM with the
received message.
Referring to \cref{fig:adderclientfsm} again,
S42 (preceding state) will be rendered when the 
endpoint is waiting for
\tmsg{RES}, but once received, S40 (successor state) 
will be immediately
rendered. The developer will need to manage the flow
of stateful application data between UI components separately
-- \fancyname{ReactMPST} only generates code for managing
the communication automaton.

@@

Receive states in \reactcodegen
work in a similar way as with \nodecodegen.
The runtime will invoke the \tmsg{RES} method
implemented by the developer upon receiving the message
through the WebSocket.
In order to do so, the S42 component 
needs to \textit{register}
the handler with the runtime.

@@

This raises questions about the type signature for the
\texttt{register} function, as the receive handler
identifiers and type signatures differ between receive states.
We adopt the same strategy as with \nodecodegen:
the receive handler registered by the receive state
to the runtime also performs the message parsing,
so all receive states register a ``general handle function'' typed

@@

\begin{figure}[!h]
\begin{lstlisting}[language=javascript,tabsize=2]
// Expect to receive `register` function passed in by runtime
type Props = {
	register: (handle: ReceiveHandler) => void
};

abstract class S42<S> extends React.Component<Props, S> {

	// Developer must implement receive handlers
	// for possible messages received at this state.
	abstract THANKS(payload1: string): void;
	abstract TERMINATE(): void;

	// When component is mounted on DOM,
	// and the register function is available,
	// register the message handler with the runtime.
	componentDidMount() {
		this.props.register(this.handle.bind(this));
	}

@@

\subsection{Terminal States}
To be consistent with the conventions defined for \nodecodegen
the browser endpoint proactively closes the connection at its terminal state.
We generate an abstract React component for terminal states 
to allow the developer to customise what to
\texttt{render} upon protocol termination.

### 2020-06-17 report/sections/server/react/limitations.tex
Report: feedback

\section{Limitations}
Our EFSM encoding provides \textbf{affine} channel
usage guarantees for send states as supposed to \textit{linear} usage
guarantees. Suppose we bind a send action to a
button click event -- strictly speaking,
even if we guarantee that the button can be clicked exactly once,
we cannot guarantee that the button will ever be clicked by the user.
However, existing work \cite{PureScript2019,MVU2020} 
share the same limitation, and we view this as an
inherent limitation in front-end development, since there is
a limit to the types of assumptions we can make about how the
user interacts with the browser.
Redefining the definition of channel linearity 
for GUI programming use cases
would be an interesting discussion.

Our runtime implementation also limits the developer from
passing data into their EFSM state component through props.
This is because the runtime takes full control of which
state component to render, and by definition, what props 
are passed to the EFSM state components.
However, we equally view this as a way to encourage developers
to decouple communication state management from
business logic state management, in the spirit
of separating concerns.
In \cref{chap:eval},
we present examples of developer implementations
that use separate abstractions (e.g. the \textit{React Context API} or
the third-party \textit{Redux} state management framework) for propagating 
stateful application data in \cref{chap:eval}.

>>>>

\section{Limitations}
\label{section:reactlimitations}

Our EFSM encoding provides \textbf{affine} channel
usage guarantees for send states as supposed to \textit{linear} usage
guarantees. Suppose we bind a send action to a
button click event -- strictly speaking,
even if we guarantee that the button can be clicked exactly once,
we cannot guarantee that the button will ever be clicked by the user.
However, existing work \cite{PureScript2019,MVU2020} 
share the same limitation, and we view this as an
inherent limitation in front-end development, since there is
a limit to the types of assumptions we can make about how the
user interacts with the browser.
Redefining the definition of channel linearity 
for GUI programming use cases
would be an interesting discussion.

Our runtime implementation also limits the developer from
passing data into their EFSM state component through props.
This is because the runtime takes full control of which
state component to render, and by definition, what props 
are passed to the EFSM state components.
However, we equally view this as a way to encourage developers
to decouple communication state management from
business logic state management, in the spirit
of separating concerns.
In \cref{chap:eval},
we present examples of developer implementations
that use separate abstractions (e.g. the \textit{React Context API}) 
for propagating 
stateful application data in \cref{chap:eval}.

### 2020-06-17 report/sections/server/react/runtime.tex
Report: complete chapter 9
Report: feedback

There are many parallels between the runtime we generate
for \fancyname{ReactMPST} compared to that of \fancyname{NodeMPST}.
We focus on the details unique to browser-side endpoints.

@@

\begin{lstlisting}[language=javascript]
type Props = {
	endpoint: string,
	states: {
		S40: Constructor<S40>,
		S42: Constructor<S42>,
		S43: Constructor<S43>,
		S41: Constructor<S41>,
	},
	waiting: React.ReactNode,
	connectFailed: React.ReactNode,
};
\end{lstlisting}

The developer provides the WebSocket URL for the runtime
to instantiate the connection.
As for the EFSM state implementations derived from the abstract
classes, the developer constructs an object to map state identifiers
to concrete implementations. 
The developer only needs to pass the \textit{constructor} function,
as the runtime will instantiate the React components separately
and pass the required props based on the type of state.
Additionally, the developer needs to define what to render
whilst \texttt{waiting} for the session to begin, and possibly
an error screen if the connection has failed.

@@

\subsection{Connecting to the Session}
We connect to the session by creating a new \texttt{WebSocket}.
Traditionally, this would be done in the constructor,
but for React components, the constructor may be invoked more
than once depending on how reconciliation works.
For this reason, we create the \texttt{WebSocket} in the
\texttt{componentDidMount} method as it is guaranteed
to be only called once.
In fact, under React's \textit{Strict Mode},
constructors are \textit{explicitly invoked twice} to 
prevent impure constructors \cite{ReactStrictMode}.

This means the WebSocket must have optional type as it is 
strictly undefined on construction.
To avoid having to deal with an optional WebSocket value
in the EFSM execution (even though we know for sure that it
has been instantiated when we use it),
we define the public API (the React component named 
\texttt{Session}) to manage the (possibly \lstonelinejs{undefined})
WebSocket in its state, and the private API (the React component
named after the role -- in this case, \trole{Client})after
as a separate non-exported class
with the WebSocket available via props. Once the WebSocket
is instantiated in the \trole{Session} component, it renders the
\trole{Client} component, passing the non-optional WebSocket value via props.

@@

The connection phase is managed in the same way as \fancyname{NodeMPST}:
when the \texttt{Client} component is mounted,
it sends a connection request to the server and overrides the
\texttt{onmessage} event handler to listen for the connection confirmation,
before advancing the EFSM to the initial state to begin executing the protocol.

\subsection{Executing the EFSM}
Unlike \fancyname{NodeMPST}, our EFSM state encoding does not define
continuation states using the actual implementation.
This is because state encodings are React components
that need to be instantiated by the runtime.
Instead, as previewed in \cref{lst:reactsendcomponent,lst:reactreceivecomponent},
we define an enum-based abstraction for specifying
successor state, and this is also used for executing the EFSM.
Our enum-based abstraction defines a string enum for each type of EFSM state,
and collects all state identifiers in an union type.

@@

We keep track of the current EFSM state in the state of the
private \texttt{Client} component.
This enables the UI to be re-rendered, and therefore ``react'',
to EFSM state transitions, which is precisely the goal of \fancyname{ReactMPST}.

@@

Messages are sent through the WebSocket in the same way 
as \fancyname{NodeMPST} -- objects marked with label and payload
are serialised into JSON. The process of sending is always followed
by a transition to the successor state, so we define the
\texttt{sendMessage()} method on the runtime to also take
the successor state as a function argument, so it can make that transition
immediately after.

@@

\subsection{Receiving Messages}

>>>>

There are many parallels between the runtime we generate
for \fancyname{ReactMPST} compared to that of \fancyname{NodeMPST}.
We focus on the details unique to React endpoints.

@@

\begin{lstlisting}[language=javascript]
type Props = {
	endpoint: string,
	states: {						// mapping of state identifier
		S40: Constructor<S40>,		// to EFSM state components
		S42: Constructor<S42>,		// implemented by developer
		S43: Constructor<S43>,
		S41: Constructor<S41>,
	},
	waiting: React.ReactNode,		// render when waiting to start
	connectFailed: React.ReactNode,	// render if connection failed
};

export class Session extends React.Component<Props, ...
\end{lstlisting}

The developer provides the WebSocket URL for the runtime
to instantiate the connection.
As for the EFSM state implementations derived from the abstract
classes, the developer constructs an object to map state identifiers
to concrete implementations. 
The developer only needs to pass the \textit{constructor} function,
as the runtime will instantiate the React components separately
and pass the required props based on the type of state.
Additionally, the developer needs to define what to render
whilst \texttt{waiting} for the session to begin, and possibly
an error screen if the connection has failed.
We show an example below.

\begin{lstlisting}[language=javascript,tabsize=2]
// Developer implementations
class InputWindow extends S40 { ... };
class QuitReceived extends S41 { ... };
class WaitScreen extends S42 { ... };
class PendingQuit extends S43 { ... };

// Main application component
class MainApp extends React.Component {
	render() {
		return <div>
			<h1>Adder Client</h1>
			<Client
				endpoint='ws://localhost:8080'
				states={{
					S40: InputWindow,
					S41: QuitReceived,
					S42: WaitScreen,
					S43: PendingQuit,		
				}}>
				waiting={<h1>Pending Connection</h1>}
				connectFailed={<p>Connection Failed</p>}
			/>
		</div>;
	}
}
\end{lstlisting}

@@

\subsection{Connecting to the Session}
We connect to the session by creating a new \texttt{WebSocket}.
Traditionally, this would be done in the constructor,
but for React components, the constructor may be invoked more
than once depending on how reconciliation works.
For this reason, we create the \texttt{WebSocket} in the
\texttt{componentDidMount}\footnote{
The \texttt{componentDidMount} method is part
of the ``commit'' phase in the lifecycle of a React component.
This is when the component is mounted on the browser DOM and
methods in the phase are guaranteed to be called exactly once.
} method as it is guaranteed
to be only called once.
In fact, under React's \textit{Strict Mode},
constructors are \textit{explicitly invoked twice} to 
prevent impure constructors \cite{ReactStrictMode}.

This means the WebSocket must have optional type as it is 
strictly undefined on construction.
To avoid having to deal with an optional WebSocket value
in the EFSM execution (even though we know for sure that it
has been instantiated when we use it),
we define the public API (the React component named 
\texttt{Session}) to manage the (possibly \lstonelinejs{undefined})
WebSocket in its state, and the private API (the React component
named after the role -- in this case, \trole{Client}) after
as a separate non-exported class
with the WebSocket available via props. 
Once the WebSocket
is instantiated in the \trole{Session} component,
it renders the
\trole{Client} component, 
passing the non-optional WebSocket value via props.

@@

The connection phase is managed in the same way as \nodecodegen:
when the \texttt{Client} component is mounted,
it sends a connection request to the server and overrides the
\texttt{onmessage} event handler to listen for 
the connection confirmation,
before advancing the EFSM to the initial state to 
begin executing the protocol.

\subsection{Executing the EFSM}
Unlike \nodecodegen, our EFSM state encoding does not define
continuation states using the actual implementation.
This is because state encodings are React components
that need to be instantiated by the runtime.
Instead, as previewed in \cref{lst:reactsendcomponent,lst:reactreceivecomponent},
we define an enum-based abstraction for specifying
successor state, and this is also used for executing the EFSM.
Our enum-based abstraction defines a string enum for each type of EFSM state,
and collects all state identifiers in an union type.

@@

We keep track of the current EFSM state in the state of the
private \texttt{Client} component.
This enables the UI to be re-rendered, and therefore ``react'',
to EFSM state transitions, which is precisely the goal of 
\reactcodegen.

@@

Messages are sent through the WebSocket in the same way 
as \nodecodegen (see \cref{subsection:noderuntimesend}) 
-- objects marked with label and payload
are serialised into JSON. The process of sending is always followed
by a transition to the successor state, so we define the
\texttt{sendMessage()} method on the runtime to also take
the successor state as a function argument, so it can make that transition
immediately after.\\

@@

\subsection{Receiving Messages}
\label{subsection:reactruntimereceive}
